---
title: "Relationship Between Black-Throated Blue Warbler Characteristics and Mass in Hubbard Brook Experimental Forest from 1986-2010"
author: "Jenna Wisniewski"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(GGally)
library(stargazer)
library(broom)
library(kableExtra)
library(naniar)
```

```{r}
#### Data Organization

## read in the data
warblers <- read_csv(here("exam_data","bt_warblers.csv"))

## convert all of the missing values stored as -99.99 to N/A
warblers_new <- warblers %>% 
  replace_with_na_all(condition =~.x == -99.99 )

```
##### **Exploratory Graphs**
```{r}
## create an exploratory ggplot of wing length vs. bird mass among different age groups with point color indicating sex and faceted by bird age at capture
ggplot(data = warblers_new, 
       aes(x = mass, y = wing_r)) + 
  geom_point(size = 2, aes(color = sex)) + 
  theme_bw() +
  facet_wrap(~ age_banded) + 
  labs(x = "Mass (grams)",
      y = "Right Wing Length (mm)", # FIXED
       title = "Warbler Wing Length vs. Mass Across Age Groups") 

```
```{r}
## create an exploratory ggplot of wing length vs. bird mass at different elevation levels with point color indicating sex and faceted by elevation of the sample plot
ggplot(data = warblers_new, 
       aes(x = mass, y = wing_r)) + 
  geom_point(size = 2, aes(color = sex)) + 
  theme_bw() +
  facet_wrap(~ elevation) + 
  labs(x = "Mass (grams)",
      y = "Right Wing Length (mm)", # FIXED
       title = " Warbler Wing Length vs. Mass Across Elevation") 
```
```{r}
## create an exploratory ggplot of bill depth vs. bill culmen among different age groups with point color indicating sex and faceted by elevation of the sample plot
ggplot(data = warblers_new, 
       aes(x = bill_culmen, y = bill_depth)) + 
  geom_point(size = 2, aes(color = sex)) + 
  theme_bw() +
  facet_wrap(~ age_banded) + 
  labs(x = "Culmen Length (mm)",
      y = "Bill Depth (mm)", # FIXED
       title = "Warbler Culmen Length vs. Bill Depth Across Age Groups") 
```


```{r, include=FALSE}
#### Multiple Linear Regression Models 

## Create a multiple linear regression model that describes the relationship between bird mass and wing length

##start by exploring the relationship between all variables using pair plots in GGally
warblers_new %>% 
  select(sex, elevation:wing_r) %>% 
  ggpairs (aes(color =sex))

```

```{r, include=FALSE}
## create multiple linear regression models in order to determine which variables yield the best fit 

lm1 <- lm(mass ~ wing_r +sex, data = warblers_new)

lm2 <- lm(mass ~ wing_r +sex +elevation, data = warblers_new)

lm3 <- lm(mass ~ wing_r +sex +elevation +age_banded, data = warblers_new)

lm4 <- lm(mass ~ wing_r +sex +elevation +age_banded + bill_depth, data = warblers_new)

lm5 <- lm(mass ~ wing_r +sex +elevation +age_banded + bill_depth + tarsus1, data = warblers_new)

lm6 <- lm(mass ~ wing_r +elevation +age_banded + bill_depth + tarsus1, data = warblers_new)

lm7 <- lm(mass ~ wing_r +sex +age_banded + bill_depth + tarsus1, data = warblers_new)

lm8 <- lm(mass ~ wing_r +sex + bill_depth + tarsus1, data = warblers_new)

lm9 <- lm(mass ~ wing_r +sex + elevation +bill_depth + tarsus1, data = warblers_new)

## Find the AIC for each model to compare and choose the lowest
AIC(lm1)
AIC(lm2)
AIC(lm3)
AIC(lm4)
AIC(lm5) 
AIC(lm7)
AIC(lm8) ## is only slightly higher than lm9
AIC(lm9) ## lowest value
```

```{r, include=FALSE}
plot(lm8) ## plotting the model to check for normality 
## appears to have relatively normal distribution and qqplot is very linear, will continue with linear regression model 
```

##### **Regression Table** 

Regression table for the relationship between warbler mass and physical variables
```{r, results = 'asis'}
## use stargazer to show the multiple linear regression coefficients for ln9
stargazer(lm8, type = "html")
```
##### **Analysis** 

The variables that I choose to use for this regression model were wing length, sex, bill depth and tarsus length (lm8). While looking at the ggpairs plots, all of these variables appeared to have relationships that predicted bird mass. Additionally, all of these variables lowered the AIC value which shows how well the model predicts the data. While elevation also lowered the AIC value, it lowered it a very small amount and when it was examined in the exploratory plots it did not appear to be a strong predictor of bird mass. Additionally, the lowering of the AIC value such a small amount was not worth the additional complexity that this variable added to the model. Banding age was not included either as the exploratory plots did not show it as a strong predictor of mass and it raised the AIC value. I did not choose bill_culmen as I had already choosen bill_depth and felt that they both would have represented bill size(their distribution on the exploratory plots was also extremely similar)and have the same prediction of mass. 

-If all other variables remain the same except for the selected variable:

-For every 1 mm increase in wing length there would be a 0.072 gram increase in mass

-Males would weigh on average 1.039 grams less than females

-For every 1 mm increase in bill depth there would be a 0.160 gram increase in mass

-For every 1 mm increase in tarsus length there would be a 0.069 gram increase in mass

##### **Citation**
Rodenhouse, N., R. Holmes, S. Sillett, and M. Webster. 2019. Black-throated Blue Warbler capture histories, Hubbard Brook Experimental Forest, 1986 - 2010 ver 4. Environmental Data Initiative. https://doi.org/10.6073/pasta/ea985022acfdb76cd7beeaf8da6c39b1
