---
title: "Changing Steller Sea Lion Pup Populations in Alaska Rookeries"
author: "Jenna Wisniewski"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(GGally)
library(stargazer)
library(broom)
library(kableExtra)
library(naniar)
```

```{r}
## Data Organization

## read in the data
steller_data <- read_csv(here("exam_data","steller_pups.csv"))

```

```{r}
## wrangle the data so that the mean value of counts is displayed by site and year
new_steller <- steller_data %>% 
  group_by(sitename, Year) %>% 
  summarize(mean_pup_count = mean(pupcount))

```

```{r}
## Create a new data frame that: 1. takes the summarized data from above and only keeps observations from sites Sugarloaf, Marmot, Forrester Complex, Chowiet, and Hazy 2. Only keeps the columns containing the site name, year, and pup count (this was already done from the previous code) 3. renames the column containing pup counts to steller_pups 

steller_counts <- new_steller %>% 
  filter (sitename %in% c("Sugarloaf","Marmot","Forrester Complex","Chowiet", "Hazy")) %>%
  rename(steller_pups = mean_pup_count)
```

##### **Sea Lion Pup Count Visualization**
```{r, fig.align = "center"}
## Create a ggplot that visualizes Steller pup population estimates over time for the five sites 
##start with a line graph
ggplot(data = steller_counts, aes(x=Year, y = steller_pups)) +
  geom_line(aes(color = sitename),show.legend = FALSE)+
##use facet wrap to show the counts by each site
  facet_wrap(~ sitename)+
##add axis labels and graph titles
  labs(x = "Year", y = "Mean Number of Pups", title = "Mean Population Count of Steller Sea Lion Pups Per 
       Alaskan Rookerie Site (1961-2015)")+
##create a background color for the graph on the axis
  theme(plot.background = element_rect(fill = "lightcyan3"))
```


**Figure 1.** Mean population count of Steller Sea Lion pups across five different Alaskan Rookeries (1961-2015) conducted by sight and aerial imaging during June and July. The Forrester Complex Rookerie recorded a 0 value for 2002 and then no further counts were recorded. Data: Fritz(2015)

##### **Data Trends**
Main trends from this data are that the Choweit, Marmot and Sugarloaf counts started off high (~5000-6000)and then all had a sharp drop in counts around the year 1980 to 1990, after which the population slowly declined until the least three years where counts have increased slightly. The Forrester Complex maintained a relatively steady count from years 1961-1998 until counts dropped to 0 in 2002 and then no further counts were recorded for that rookerie. In 1979 the Hazy Rookerie counts began and have generally increased slowly over time. 

##### **Citation** 
Fritz, Lowell; Sweeney, Kathryn; Lynn, Morgan; Gelatt, Thomas; Gilpatrick, James; Towell, Rodney (2015). Counts of Alaska Steller sea lion pups conducted on rookeries in Alaska from 1961-06-22 to 2015-07-18 (NCEI Accession 0128189). NOAA National Centers for Environmental Information. Dataset. https://doi.org/10.7289/v5862ddr. Accessed 2020-12-04.