---
title: "Sea Turtles in Shark Bay Data"
author: "Jenna Wisniewski"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(GGally)
library(stargazer)
library(broom)
library(kableExtra)
library(naniar)
library(effsize)
```

```{r}
## Data Organization

## read in the data
turtles_data <- read_csv(here("exam_data","sharkbay_sea_turtles.csv"))
```
#### **1. Explore the relationship between carapace length and width for green and loggerhead turtles**


##### **Data Vizualization**
```{r}
## create an exploratory data visualization showing the relationship between carapace length and width 

ggplot(data = turtles_data, aes(x=length, y = width)) +
  geom_point()+
  labs(x = "Carapace Length(cm)", y = "Carapace Width (cm)", title = "Relationship Between Carapace Length and Width in Green and Loggerhead Turtles")

```


##### **Analysis**

```{r, include=FALSE}
## Do linear regression do get the R-squared value
turtles_lm <- lm(length ~ width, data = turtles_data)

## check diagnostic plot to check assumptions
plot(turtles_lm)
## Data looks like it meets assumption, however might possibly violate linearity and could have outliers that are affecting the data.

# Get tidy versions of the model output to call later on in text
turtles_lm_tidy <- tidy(turtles_lm)

turtles_lm_glance <- glance(turtles_lm)
```

 Simple linear regression revealed that carapace length strongly predicts carapace width in green and loggerhead turtles, with 93% of the variance in carapace width accounted for by carapace length p < 0.005, R^2^ = `r round(turtles_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(turtles_lm_tidy$estimate[2], 2)` cm (i.e., for each one centimeter increase in carapce length we expect an average increase in carapce width of `r round(turtles_lm_tidy$estimate[2], 2)` cm). Diagnostic plots (not included) reveal normally distributed and homoscedastic residuals. 

#### **2. Compare the carapace lengths between green and loggerhead turtles**
```{r}
## Data Organization

## create a new data set that just contains species and carapace length.

turtles_length <- turtles_data %>% 
  select("species", "length")

## Create vectors to make exploratory plots

loggerhead_l <- turtles_length %>% 
  filter(species == "loggerhead") %>% 
  pull(length)

green_l <- turtles_length %>% 
  filter(species == "green") %>% 
  pull(length)

```
##### **Data Visualization: Exploratory Plots**
```{r}
## Create exploratory plots to explore the distribution of carapace lengths for green and loggerhead turtles 

##Plot data in exploratory graph 
hist(loggerhead_l) ## Looks relatively normal possibly a bit skewed right
qqnorm(loggerhead_l) ## looks very linear
hist(green_l) ## Looks very normal 
qqnorm(green_l)## Looks relatively linear, possibly a bit skewed

##Overall: assumptions of normality hold, & sample size additionally >> Central Limit Theorem (means will be normally distributed). Two-sample t-test is okay for means comparison and will be used
```

##### **Analysis**

To compare the lengths statistically, I choose a two sample t test. This is because


```{r, include=FALSE}
##Use two sided t.test to compare the mean carapace lengths of loggerhead and green turtles
turtles_ttest <- t.test(x = loggerhead_l, y = green_l)

turtles_ttest_tidy <- tidy(turtles_ttest)
```

```{r, include=FALSE}
 # Get the means & SDs for each: 
loggerhead_mean <- mean(loggerhead_l,  na.rm = T)
green_mean <- mean(green_l)
loggerhead_sd <- sd(loggerhead_l, na.rm = T)
green_sd <- sd(green_l)
```

```{r, include=FALSE}
## Completing a cohen.d test to examine the difference between the variance of points
turtles_effsize <- cohen.d(loggerhead_l, green_l, na.rm = T)
```
On average, green turtles have a longer carapace than loggerhead turtles. (`r round(green_mean,2)` $\pm$ `r round(green_sd, 2)` and `r round(loggerhead_mean,2)` $\pm$ `r round(loggerhead_sd, 2)` cm, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(green_mean, 2) - round(loggerhead_mean, 2)` cm (a `r round((green_mean - loggerhead_mean)/((green_mean + loggerhead_mean)/2)*100, 2)`% difference), the difference in means is not significant (Welch's two-sample t-test: t(`r round(turtles_ttest_tidy$parameter,2)`) = `r round(turtles_ttest_tidy$statistic,2)`,  p = (`r format(turtles_ttest_tidy$p.value, scientific = TRUE, digits = 3)`) , p is not < 0.05), and the effect size is small (Cohen's *d* = `r round(turtles_effsize$estimate,2)`).


#### **3. Comparing the Occurrence of burrowing barnacles on green vs. loggerhead turtles**
```{r}
#### Data Organization

##Create a subset of data that only selects species and burr (presence of barnacles) 

turtles_b <- turtles_data %>% 
  select(species, burr) %>% 
## change the yes and no responses to "with burrowing barnacles observed" and "without burrowing branacles observed"
 mutate(burr = case_when(
 burr == "yes" ~ "Burrowing Barnacles Observed",
 burr == "no" ~ "Burrowing Barnacles Not Observed")) 

        
```

##### **Data Visualization**


**Table 1.** Counts of green and loggerhead turtles that are observed with the presence of burrowing barnacles in Shark Bay Australia during at-sea surveys. Data: Heithaus (2019)
```{r}
##creating a table that shows the counts from each response in correlation to the identified political affiliation 
turtles_counts <- turtles_b %>% 
  janitor::tabyl(species, burr) 

turtles_table <- turtles_counts %>% 
  adorn_percentages() %>% 
   janitor::adorn_pct_formatting(digits = 2) %>% 
  adorn_ns() 

 turtles_table %>% 
   kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)##finalize the graph format 

```
##### **Analysis**

```{r, include=FALSE}
##converting count table into contingency table so that I can perform a chi square test

turtles_ct <- turtles_counts %>% 
  column_to_rownames(var = "species")
```

```{r, include=FALSE}
## to address the question- is there a significant association between species and occurrence of burrowing barnacles run a chi squared test to test the affiliation
turtles_x2 <- chisq.test(turtles_ct)
turtles_x2

turtlesx2_tidy <- tidy(turtles_x2)

turtlesx2_tidy
```

The percentage and count of green turtles with observed burrowing barnacles was `r turtles_table[1,2]`, the percentage and count of loggerhead turtles with observed burrowing barnacles was `r turtles_table [2,2]`.  The chi squared value from the two species of `r round(turtlesx2_tidy$statistic,1)` and a p value of `r format(turtlesx2_tidy$p.value, scientific = TRUE, digits = 3)`, p < 0.05 is sufficient evidence to reject the null hypothesis of independence between turtle species and presence of burrowing barnacle, showing that the variables are dependent on each other and turtle species will effect whether burrowing barnacle is observed an a trapped individual. 

